Architectures for workflow integration in science gateways

### 2016-04-08 paper.tex
Added figures, updates in paper.tex

\documentclass[11pt]{article}

\title{Architectures for worklfow integration in science gateways}

@@

\begin{document}

@@

Goal: review and compare the different architectures used to integrate
workflows in science gateways, based on practical experience.  A
workflow engine is defined as a software that submits jobs to a
cluster, grid or cloud. Workflows may be expressed in any language,
including scripts. Several workflow engines may be combined in the
same science gateway.

Context: The question of integrating workflow engines in science gateways can
be seen at various levels, corresponding to various definitions of
workflows. One level is the SHIWA level, where it was considered that
workflow engines are aware of the DCI (and 'D' is important because of
data transfers, proxies, etc). Another level is to remove 'D' from the
definition: a workflow engine becomes a program that submits jobs,
potentially only to local clusters. It opens a whole new class of
workflow engines that we used to consider as "applications". For
instance, in neuroinformatics: Nipype, PSOM, but also FSL through the
fslsub tool. A workflow engine is not supposed to be aware of the science gateway.

Without loss of generality, several of the examples presented in the
paper will be taken from medical image analysis, in particular
neuroimaging.

SG: a web portal with some features to handle DCIs (data transfers,
job submission, auth, etc). An ecosystem of services to access
distributed infrastructures. Examples: CBRAIN, VIP, NSG, P-GRADE, etc

Infrastructure: computing and storage resources. Includes schedulers, catalogues, etc.

Workflow: a process that submits tasks to the infrastructure. FSL
tools, say Feat, can be considered as workflows when they are
configured to submit tasks to clusters using fslsub.

A workflow consists of activities.

Contribution of the paper:
\begin{itemize}
\item We evaluate architectures based on our experience with existing systems
\item We propose a new architecture
\end{itemize}

We are talking only about concrete workflows with a configuration (see terstyansky et al 2014 "enabling scientific workflow sharing...")

@@

@@

@@

@@

@@

@@

\bibliographystyle{plain} 
\bibliography{biblio}

\end{document}

>>>>

\documentclass[preprint,3p,twocolumn]{elsarticle}

\usepackage{svg}

\newcommand{\todo}[1]{\color{blue} #1 \color{black}}
\newcommand{\note}[1]{\color{blue}\textbf{Note: }\textit{#1} \color{black}}

\begin{document}
\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Architectures for workflow integration in science gateways}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{}

\address{}

\begin{abstract}
%% Text of abstract

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

@@

\journal{Future Generation Computer Systems}

@@

\paragraph{Context} The question of integrating workflow engines in
science gateways can be seen at various levels, corresponding to
various definitions of workflows. One level is the SHIWA level, where
it was considered that workflow engines are aware of the DCI (and 'D'
is important because of data transfers, proxies, etc). Another level
is to remove 'D' from the definition: a workflow engine becomes a
program that submits jobs, potentially only to local clusters. It
opens a whole new class of workflow engines that we used to consider
as "applications". For instance, in neuroinformatics: Nipype, PSOM,
but also FSL through the fslsub tool. A workflow engine is not
supposed to be aware of the science gateway.  A wide-array of workflow
engines are available with specificies coming from the application
domain, available tools, etc Their integration in science gateways
becomes critical. Several of the examples presented in the
paper will be taken from medical image analysis, in particular
neuroimaging.

\paragraph{Goal} this paper reviews and compares the architectures to
integrate workflow engines in science gateways.

\paragraph{Contributions}
\begin{itemize}
\item We evaluate architectures based on our experience with existing systems
\item We propose a new architecture
\end{itemize}

\section{Workflow engines and science gateways}

\subsection{Workflow engines}

In the last decade, the e-Science workflow community has developed
high-level workflow systems to help developers access distributed
infrastructures such as grids and web services, resulting in tools
such as Askalon, Hyperflow, MOTEUR, Pegasus, Swift, Taverna, Triana,
VizTrails, WS-PGRADE/gUSE, etc. Such workflow engines usually describe
applications in a specific language that offers operators for parallel
computing, visual description and edition, links with domain-specific
tool repositories, etc. Descriptions and experiments conducted with
e-Science workflow engines were published in journals such as Future
Generation Computer Systems and conference venues such as
WORKS.

At the same time, specific toolboxes were emerging in various
scientific domains to facilitate interactions between different
software components. In neuroimaging for instance, tools such as
Nipype, PSOM, SPM and FSL provide abstractions and functions to define
processes that handle the data flow between other processes. Soon,
such tools were interfaced to computing systems, in particular
clusters: they were extended to create cluster tasks, handle their
dependencies, and execute them on clusters. For instance, FSL can
launch tasks on SGE through its \texttt{fsl\_sub} tool, Nipype
\todo{...}, PSOM \todo{...}, and SPM \todo{...}. Although distributed
infrastructures are hardly handled, such domain-specific tools have to
be called workflow engines. They represent a tremendous opportunity
for science gateways to leverage existing tools and
applications. Whether these should be combined with e-Science workflow
engines is an open question.

A workflow engine is defined as a software that submits jobs to a
cluster, grid or cloud. Workflows may be expressed in any language,
including scripts. Several workflow engines may be combined in the
same science gateway. Workflow: a process that submits tasks to the
infrastructure. FSL tools, say Feat, can be considered as workflows
when they are configured to submit tasks to clusters using fslsub. A
workflow consists of activities. We are talking only about concrete
workflows with a configuration (see terstyansky et al 2014 "enabling
scientific workflow sharing...")

\subsection{Science gateways}

Science Gateway: An ecosystem, usually accessible through a web
  portal, that provides tools to access distributed
  infrastructures. Tools usually help users manage data transfers,
  task execution and authentication on multiple computing and storage
  locations. Examples: CBRAIN, VIP, NSG, etc

\subsection{Infrastructures}

\todo{Maybe this is not relevant}

@@

We define architectures from their main components and the interactions
between these components.

A workflow engine is not supposed to be aware of the science gateway (?).

\subsection{Common components}

Infrastructure: computing and storage resources. Includes
  schedulers, catalogues, etc. Can be grid, cloud or even
  cluster. Some workflow engines may assume a shared file system
  usually available only on clusters.

  Workflow engine: executes workflows, i.e. process their dependencies
  and create computing tasks. Might transfer data too.

\subsection{Interactions}

a.

d. Sub-tasking: science gateway knows the sub-tasks of a
task. Therefore, it can kill all sub-tasks of a task when a task is killed.

g. Workflow integration results in an interface, for instance a web form,
where users can enter the parameters of the workflow to be
executed. Integrating a workflow is not the same process as
integrating a workflow engine. Integrating a workflow engine results
in an interface that allows developers to enter the parameters of the
workflow engine.

Only workflow execution is described. 
Workflow integration is not covered.

@@

\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/tight.pdf_tex}
\label{archi:tight}
\caption{Tight integration}
\end{figure}

@@

\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/service.pdf_tex}
\label{archi:service}
\caption{Service integration}
\end{figure}

@@

No global vision of the workflow.

Science gateway has to make sure that the machine on which a workflow
task executes is reliable as several sub-tasks will depend on the
success of this task. Or workflow engine task has to be recoverable.

Performance: there could be a deadlock when there is contention. Also,
the workflow task has to be reliable (for instance, proper walltime
estimation), which is difficult because these tasks are
long. Restarting a workflow task may mean that all the tasks in the
workflow have to be restarted, depending on the capability of the
workflow engine. This complexifies task scheduling (a system could
ignore these issues but they may backfire).

@@

These architectures can be combined (give examples).

@@

\bibliographystyle{elsarticle-num} 
\bibliography{biblio}

\end{document}
\endinput
